---
title: "Welcome to R-Ladies St. Louis"
author: "Introduction to Time Series"
date: "April 13, 2021"
output:
  xaringan::moon_reader:
    css: ["default", "rladies", "rladies-fonts"]
    nature: 
        highlightLines: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Materials for tonight 

<br>

- All materials can be found here: 
https://github.com/rladiesstl/meetup-apr-2021

- Install R packages:

  + feasts
  + fable
  + tsibble
  + tidyverse
  + readxl
  + here


---

# What is R-Ladies?

- R-Ladies is a world-wide organization with the mission of promoting **gender diversity in the R community**  

- Started in San Francisco in 2012 and now has **205 chapters and more than 86,000** members globally (check out rladies.org for a Shiny dashboard)

```{r, echo=FALSE, fig.align="center", out.width='70%'}
knitr::include_graphics("https://pbs.twimg.com/media/EPlDKUQUcAEq5pw?format=jpg&name=small")

```

---

# R-Ladies St. Louis 

* Started in September 2017 by Jenine Harris and Chelsea West 

* 4 Meetups in 2017-2018, 5 Meetups in 2018-2019, 5 in 2019-2020, and 7 in 2020-21!

* Co-organizers for 2020-21 include:

  - Jenine Harris (Washington University)  
    
  - Mary Painter (Univeristy of Missouri)  
    
  - Shelly Cooper (Washington University)  
    
  - Crystal Lewis (University of Missouri)
  
---
  
# 2020-21 Schedule

* Tonight is our last meetup of the season!
    
* RSVP for future talks on meetup.com

* We will be recording this talk
  
    - Feel free to turn off your cameras if you do not want to be recorded
    - Use the chat to ask questions (to everyone **or** directly to the host)

---

# Want to learn more or get involved?

- Visit **rladies.org** to learn about the mission and global work of R-Ladies 

- Connect with R-Ladies St. Louis 
    + Email us at <a href = "mailto:rladiesstl@gmail.com">rladiesstl@gmail.com</a> 
    + Follow us on Twitter <a href="https://twitter.com/RLadiesSTL">@rladiesstl</a>
    + Connect with us on <a href="https://www.meetup.com/R-Ladies-St-Louis/">Meetup</a>
    
<img src = "https://github.com/rladies/starter-kit/blob/master/logo/R-LadiesGlobal_RBG_online_LogoOnly.png?raw=true" width = 250/><img src = "https://media.giphy.com/media/3o7abBphHJngINCHio/giphy.gif" width = "300" align = "right"/>

---

# About tonight's speaker

- Laura Rose works for Mallinckrodt Pharmaceuticals

- She is a Manager of Demand where she forecasts Active Pharmaceutical Ingredients

- She has an M.A. in Economics from the University of Missouri-St. Louis

- She enjoys using R for time series analysis, building Shiny apps, and automating data analysis tasks


